# PA1_template.rmd


#### set the global options
```{r, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

#### get the source data for processing. data is stored locally
```{r loaddata, chache = TRUE}
setwd("/Users/boopathi/Datascience/R/assignments/rmarkdown")
rawdata <- read.csv("activity.csv")
```

#### check the data 
```{r checkdata}
summary(rawdata)
dim(rawdata)
str(rawdata)
```

#### clean the data
```{r cleandata}
activity <- rawdata #for now dont remove NAs
activity$date <- as.Date(activity$date)
```

#### summarize the data
```{r summarizeData, echo=FALSE, results='hide'}
library(dplyr)
actsum <- rawdata %>%
    group_by(date) %>%
    summarise(totalSteps = sum(steps))
```

#### plot the histogram
```{r histofsteps}
hist(actsum$totalSteps, xlab = "Total Steps", main = "Total number of steps per day")
```

#### mean and median from Summary
```{r meanandmedian}
summary(actsum$totalSteps)
```

#### avg number of stpes based on time interval and max of avg
```{r avgsteps}
actsum2 <- rawdata %>% group_by(interval) %>%
    summarise(avgSteps = mean(na.omit(steps)))
plot(actsum2$interval, actsum2$avgSteps, type = 'l', xlab = "Interval", ylab = "avg steps")
maxval <- actsum2[which(actsum2$avgSteps == max(actsum2$avgSteps)),]
points(maxval, type="p", pch = 19, col = "blue")
text(maxval, labels = paste0("max", ": ", maxval[,1], ",", round(maxval[2])), pos = 4)
```

#### imupting data
##### using mean of interval
```{r imuptingData}
rawdata$tempint <- as.character(rawdata$interval)
actsum3 <- rawdata %>% 
    group_by(tempint) %>%
    summarise(meanSteps = as.integer(mean(na.omit(steps)))) #calcualtes mean by intervals
rawdata <- merge(rawdata, actsum3, by.x = "interval", by.y = "tempint") #add new Steps to rawdata
rawdata$newSteps <- if_else(is.na(rawdata$steps), as.integer(rawdata$meanSteps), rawdata$steps) #use mean where stpes is NA
```


#### summarize the data, with the imputed data
```{r summarizeDataNew, echo=FALSE, results='hide'}
library(dplyr)
actsum4 <- rawdata %>%
    group_by(date) %>%
    summarise(totalStepsNew = sum(na.omit(newSteps)))
```

#### plot the histogram
```{r histofstepsNew}
hist(actsum4$totalStepsNew, xlab = "Total Steps", main = "Total number of steps per day - imputed")
```

#### mean and median total number of steps taken per day
```{r meanandmediannew}
#before imputing
summary(actsum$totalSteps)
#after imputing
summary(actsum4$totalStepsNew)
```

#### activity paterns by weekdays vs weekends
```{r weekdayvsend}
rawdata$date <- as.Date(rawdata$date)
rawdata$day <- weekdays(rawdata$date, abbreviate = T)
rawdata$weekdayind <- as.factor(ifelse(rawdata$day %in% c("Sat","Sun"), "Weekend","Weekday"))

actsum5 <- rawdata %>% 
    group_by(weekdayind,interval) %>%
    summarise(meanNewSteps = mean(newSteps))

library(ggplot2)

p <- ggplot(data = actsum5, aes(interval, meanNewSteps)) #base plot
p <- p + geom_line() # add line
p <- p + facet_grid(weekdayind ~.) # add facets
p <- p + xlab("Interval") + ylab("Average Steps") 
plot(p)
```




